Uncovering Distributed System Bugs\\ during Testing (not Production!) -- draft title

### 2015-08-18 paper.tex
High-level modeling approach for migration.  I haven't added the figure or explained async yet.

We also used \psharp to test the MigratingTable library, which can be integrated into an application that accesses data using the Windows Azure table storage service and is capable of migrating the data from one underlying Azure table to another, transparent to the application.  MigratingTable provides a ``virtual table'' with an API similar to that of an ordinary Azure table.  Each read and write on the virtual table is translated to a sequence of reads and writes on the underlying tables according to a protocol we designed, which guarantees linearizability of operations on the virtual table across multiple application processes assuming that the underlying tables respect their own linearizability guarantees.

@@

\subsubsection{Modeling approach}
% TBD where this goes ends up in the paper ~ Matt 2015-08-17

TODO

>>>>

We also used \psharp to test the MigratingTable library, which is capable of transparently migrating a data set between tables in the Windows Azure storage service while an application is accessing it.  MigratingTable provides a ``virtual table'' with an API similar to that of an ordinary Azure table, backed by a pair of ``old'' and ``new''  tables.  It moves all data from the old table to the new table in the background.  Meanwhile, each read or write issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol we designed, which guarantees linearizability of operations on the virtual table across multiple application processes assuming that the backend tables respect their own linearizability guarantees.

@@

\subsubsection{Modeling approach}
% TBD where this ends up in the paper ~ Matt 2015-08-17

MigratingTable implements an interface called IChainTable2, which provides the core read and write functionality of the real Azure table API with one exception: it provides a weaker consistency property for multi-page reads, since the original property would have been difficult to achieve for no benefit to applications we could foresee.  MigratingTable requires that its backend tables also implement IChainTable2, and we wrote a simple adapter to expose physical Azure tables as IChainTable2.  Our goal was then to verify that when multiple application processes issue ``input'' read and write calls to their own MigratingTable instances backed by the same tables, the behavior is compliant with the specification of IChainTable2 for the combined input history.

% N.B. SpecTable = InMemoryTableWithHistory in the current codebase. ~ Matt 2015-08-17
Since the specification is deterministic under sequential calls except for the results of multi-page reads, we decided the easiest way to formulate it for automated testing was to write an in-memory reference implementation called SpecTable.  Given a multi-page read, SpecTable can actually produce a list of all valid results.  Our correctness property is then:
\begin{quote}
For every execution trace of a collection of MigratingTables backed by the same pair of SpecTables (which nondeterministically choose one of the valid results for each multi-page read), there exists a linearization of the combined input history such that the output in the original trace matches the output of a ``reference'' SpecTable on the linearized input.
\end{quote}
%
\def\term#1{\emph{#1}}
We instrumented MigratingTable to report the \term{linearization point} of each input call, which in our case is always one of the corresponding \term{backend calls} to the backend tables (often the last).  Specifically, after each backend call, MigratingTable reports whether it was the linearization point, which may depend on the result of the call.  This makes it possible to verify the correctness property as the system executes.  We have a \term{tables machine} containing all three SpecTables and a collection of \term{service machines} each containing a MigratingTable.  Each service machine issues a random sequence of input calls to its MigratingTable, which sends backend calls to the tables machine.  When MigratingTable reports the linearization point of an input call, the service machine sends that input call to the reference table.  When an input call completes, the service machine checks that the results from the MigratingTable and the reference table agree.  \psharp controls the interleaving of the backend calls.  After the tables machine processes a backend call, it enters a state that defers further backend calls until MigratingTable has reported whether the backend call was a linearization point and (if so) the call to the reference table has been made.  We also use the \psharp nondeterminism API to generate the input calls, so in principle an exhaustive \psharp scheduling strategy such as DFS could be used to exhaustively test MigratingTable up to some bound.

To do...
